---
title: "PracticalMachineLearning_RMD1"
author: "Sachinder"
date: "Wednesday, June 17, 2015"
output: html_document
---
##Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal is to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har.

## Loading & Tidying data
The training data for this project are available here and the tests data here 
After downloading into the working directory, the following code loads them in training and test. Needed libraries are also loaded.

```{r}

library(randomForest)
library(caret)
library(parallel)
library(doParallel)
setwd("C:/Users/sachinderk/Documents/PracMachine")
#load raw data
trainingSrc   <- read.csv('pml-training.csv', na.strings=c("NA","#DIV/0!", ""))
testSrc       <- read.csv('pml-testing.csv' , na.strings=c("NA", "#DIV/0!", ""))
```

We decided to leave columns having no more than 60% of NA values:
```{r}
goodVars    <- which((colSums(!is.na(trainingSrc)) >= 0.6*nrow(trainingSrc)))
trainingSrc <- trainingSrc[,goodVars]
testSrc     <- testSrc[,goodVars]
```

Some minor fixes to test set are needed to perform well with random forests.
```{r}
# remove problem id
testSrc <- testSrc[-ncol(testSrc)]
# fix factor levels
testSrc$new_window <- factor(testSrc$new_window, levels=c("no","yes"))
```
Remove X and cvtd_timestamp colums from the dataset since they are not relevant
```{r}
trainingSrc <- trainingSrc[,-c(1,5)]
testSrc     <- testSrc[,-c(1,5)]
```

Splitting data
Split the data: 60% for training, 40% for testing
```{r}
inTraining  <- createDataPartition(trainingSrc$classe, p = 0.6, list = FALSE)
training    <- trainingSrc[inTraining, ]
testing     <- trainingSrc[-inTraining, ]

```

We create train and testing sets to improve the quality of the model.To reduce the dimension of the data I used PCA. I have chosen Parallel Random Forest as my modeling method. Note that for random forests there is no need for cross-validation to get an unbiased estimate of the test set error. It is estimated internally during the fitting process.

##Firrting Random Forest

The outcome variable is class and other colums are in data dataframe.
```{r}
class <- training$classe
data  <- training[-ncol(training)]
registerDoParallel()
rf <- train(data, class, method="parRF", 
    tuneGrid=data.frame(mtry=3), 
    trControl=trainControl(method="none"))
```
```{r}
rf
```
Let's plot importance of the model variables:
```{r}
plot(varImp(rf))
```

##Confusion Matrix for testing set

Predict on testing set and generate the confusion matrix for the testing set

```{r}
testingPredictions <- predict(rf, newdata=testing)
confMatrix <- confusionMatrix(testingPredictions,testing$classe)
confMatrix
```
Let's have a look at the accuracy
```{r}
confMatrix$overall[1]
```
It looks very good - it is more then 99,5%.

##Submit results of Test Set
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

answers <- predict(rf, testSrc)
pml_write_files(answers)
```


